---
title: "Calicorema Capitata Surface Soil Water Potentials"
author: "Avery Swanton, Daisy Amaro, April Ibarra, Ben H, Isabel Torres"
date: "7/1/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

All soil samples were collected at the surface level. 
Treatment types are: -DN: dune nara -RO: river open
                    -DO: dune open -RN: river open


```{r} 
library(lme4)
library(ggplot2)
library(cowplot)
library(gplots)
library(tidyverse)
```

```{r P data import and summary}
data <- read.csv("https://www.dropbox.com/scl/fi/v87x0dijgyuluq70l4mm4/swp.csv?rlkey=h9i500avqxb38gxklydi83274&st=ghmfv5sf&dl=1")
str(data)
head(data)
data$Location<- as.factor(data$Location)
data$Treatment<- as.factor(data$Treatment)
data$Date<- strptime(data$Date, format= "%m/%d/%y") #parses date into a date that R understands
data$Date<- as.POSIXct(data$Date)

```


```{r}
SWP.model<- lm(SWP..MPa.~Location*Treatment, data = data)
SWP.model

anova(SWP.model)

```

#practice for 3 variable time series -works but not seprated by groups
```{r}
ggplot(data = data, aes(x = Date, y = SWP..MPa.)) +
  geom_point(color = "red", lty = 1) +
  theme_bw()


```


#works
```{r}
SWP.model2<- lm(SWP..MPa.~Treatment*Location, data = data)
SWP.model2

anova(SWP.model2)

```


```{r}
ggplot(data = data, aes(x = Date, y = SWP..MPa.)) +
  geom_line(color = "red", lty = 1) +
  theme_bw()


```


#boxplot 
```{r}
boxplot(data$SWP..MPa.~Treatment,data=data, main="Soil Water Potential of surface soils",
   xlab="Treatment Type", ylab="Soil Water Potential (Mpa)")

boxplot(data$SWP..MPa.~data$Treatment * Date,data=data, col=(c("red","orange","blue")), main="Soil Water Potential of surface soils",
   xlab="Treatment and Date", ylab="Soil Water Potential (MPa)")
legend(x = -800, y = 5, # Coordinates
       legend = c("Control", "Treatment 1", "Treatment 2"),
       lty = c(1, 2, 3),
       col = c(4, 6, 3),
       lwd = 3)
```

```{r}
#For Control soils
data_control <- subset(data, data$Treatment  == "Control", )
data_control


linear_regression_plot <- ggplot(data = data_control, aes(x = SWP..MPa., y = Treatment)) +
  geom_point(size=2, shape=1) +
  theme_bw()+ geom_smooth(method=lm)
linear_regression_plot

linear_model_SWP_treatment_type<- lm(SWP..MPa. ~ Date*Location , data = data_control)
summary(linear_model_SWP_treatment_type)

plot(linear_model_SWP_treatment_type)

boxplot(SWP..MPa.~Date,data=data_control, main="Soil Water Potential of Control Surface Soils",
   xlab="Date", ylab="Soil Water Potential (MPa)")
#calculate averages for C soils 
c.avg<- mean(data_control$SWP..MPa.)

```

```{r}
#For Treatment 1 soils
data_T1 <- subset(data, data$Treatment  == "SmallPulse", )
data_T1


linear_regression_plot <- ggplot(data = data_control, aes(x = SWP..MPa., y = Treatment)) +
  geom_point(size=2, shape=1) +
  theme_bw()+ geom_smooth(method=lm)
linear_regression_plot

linear_model_SWP_treatmentT1_type<- lm(SWP..MPa. ~ Date*Location , data = data_T1)
summary(linear_model_SWP_treatmentT1_type)

plot(linear_model_SWP_treatmentT1_type)

boxplot(SWP..MPa.~Date,data=data_T1, main="Soil Water Potential of Small Pulse Surface Soils",
   xlab="Date", ylab="Soil Water Potential (MPa)")
#calculate averages for T1 soils 

```

```{r}
#For Treatment 2 soils
data_T2 <- subset(data, data$Treatment  == "BigPulse", )
data_T2


linear_regression_plot <- ggplot(data = data_control, aes(x = SWP..MPa., y = Treatment)) +
  geom_point(size=2, shape=1) +
  theme_bw()+ geom_smooth(method=lm)
linear_regression_plot

linear_model_SWP_treatmentT2_type<- lm(SWP..MPa. ~ Date*Location , data = data_T2)
summary(linear_model_SWP_treatmentT2_type)

plot(linear_model_SWP_treatmentT2_type)

boxplot(SWP..MPa.~Date,data=data_T2, main="Soil Water Potential of Big Pulse Surface Soils",
   xlab="Date", ylab="Soil Water Potential (MPa)")
```

#geom ine, separate treatments over time, need IQR and error bars, using just the means. 
```{r}
CPlot <- ggplot(data = data_control, aes(x = Date, y = SWP..MPa.))+ geom_line(color = "red", lty = 1) +
  theme_bw()
T1Plot <- ggplot(data = data_T1, aes(x = Date, y = SWP..MPa.))+ geom_line(color = "orange", lty = 1) +
  theme_bw()
T2Plot <- ggplot(data = data_T2, aes(x = Date, y = SWP..MPa.))+ geom_line(color = "blue", lty = 1) +
  theme_bw()

multiplot<- plot_grid(CPlot, T1Plot, T2Plot, labels = c('Control', 'Treatment 1', 'Treatment 2'))
multiplot 
ggsave(multiplot,file = "multiplot.jpeg", width=15, height = 15, units = "in")
```

#07/20 works
```{r}
SWPPlot <- ggplot(data = data, aes(x = Date, y = SWP..MPa.)) + geom_point(aes(color = Treatment, shape= Treatment))
  theme_bw()
  SWPPlot
  
LPlot <- ggplot(data = data, aes(x = Date, y = SWP..MPa.)) + geom_point(aes(color = Location, shape= Treatment))
  theme_bw()
  LPlot

ggsave(SWPPlot,file = "multiplot.jpeg", width=15, height = 15, units = "in")

#cali fake data r script analysis on usb 
```
```{r}
boxplot(SWP..MPa.~Date,data=data, main="Soil Water Potential of Treatment 2 Surface Soils",
   xlab="Date", ylab="Soil Water Potential (Mpa)")


```
#from heather's code 
# Reduce to one mean value for treatment per date -----------------------------------------------------
```{r}
SWP_trt_means <- data |> 
  group_by(Date, Treatment) |> 
  summarize(
    N    = sum(!is.na(SWP..MPa.)),
    mean.SWP = mean(SWP..MPa.),
    sd_SWP   = sd(SWP..MPa., na.rm=TRUE), # not working at the moment...
    se_SWP   = sd_SWP / sqrt(N) 
  )

# Quick plot -----------------------------------------------------

SWP_plot<- SWP_trt_means |>
  ggplot(aes(x = Date, y = mean.SWP)) +
  geom_point(aes(color=Treatment)) +
 geom_errorbar(aes(ymin=mean.SWP-se_SWP, ymax=mean.SWP+se_SWP)) +
  geom_line(aes(color = Treatment))
SWP_plot
```

#try again for treatment, location seprate graphs per location

